---
title: "01 - Dataset Audit"
subtitle: "Leakage-Aware SMO/SVM Transcriptomic Biomarker Classification"
author: "Project team"
format:
  html:
    toc: true
    code-fold: false
execute:
  warning: true
  message: true
---

## Purpose

This notebook documents the **dataset audit** (Table 1) for the project.
It is the explanatory companion to `scripts/01_dataset_audit.R`.

Per `CLAUDE.md`, this step does **not** download or commit large raw
expression matrices. We record **only** accession-level metadata and
cohort summaries. All numeric cohort counts start as `NA` placeholders
and **must be verified by a human** against the official GEO records
before any downstream use.

## Scope of the audit

Cohorts considered:

- **GSE25055** - discovery (model development / nested CV).
- **GSE25065** - external validation.
- **GSE20194** - external validation (MAQC-II; high overlap risk).
- **GSE41998** - optional, pending independence check.
- **GSE20271** - optional, pending independence check.

Endpoint of interest: pathologic complete response (**pCR**) vs
residual disease (**RD**) after neoadjuvant chemotherapy.

## Leakage-control reminders

Following `CLAUDE.md` (scientific rules):

1. External validation cohorts must never inform feature selection,
   threshold tuning, hyperparameter tuning, or batch correction.
2. All supervised steps belong inside training folds only.
3. Sample overlap across MAQC-II-derived series (GSE25055 / GSE25065 /
   GSE20194) is a concrete leakage risk and must be quantified at the
   GSM level before these cohorts are combined.

## Build the audit table

The actual table construction lives in the script so it stays
reproducible. Here we simply source it and display the result.

```{r}
#| label: run-audit
source(here::here("scripts", "01_dataset_audit.R"))
```

```{r}
#| label: show-table
audit_path <- here::here("tables", "table1_dataset_audit_filled.csv")
audit_tbl <- readr::read_csv(audit_path, show_col_types = FALSE)
knitr::kable(audit_tbl)
```

## Human-verification checklist

> The values below are **not** final. A human must confirm each item
> against the GEO series pages and study supplements.

- [ ] Confirm platform (GPL) accession for each series.
- [ ] Fill `total_n`, `pcr_n`, `rd_n` from curated clinical annotation
      (not the raw GSM count).
- [ ] Confirm the exact pCR / RD endpoint definition per study.
- [ ] Cross-check GSM identifiers to quantify true sample overlap.
- [ ] Decide inclusion of optional cohorts only after independence is
      established.

## Output

Running this notebook (or the script) writes:

- `tables/table1_dataset_audit_filled.csv`

No raw expression data is written to the repository.
