---
title: "Pilot: Leaky vs Nested SMO/SVM on GSE25055"
subtitle: "First reproducible pilot (discovery cohort only)"
format: html
execute:
  echo: true
  warning: true
  message: true
---

# Purpose

This pilot compares two transcriptomic classification pipelines on **GSE25055
only**:

- **Pipeline A (naive / leaky baseline):** supervised feature selection is
  performed **globally, before** cross-validation. This deliberately leaks
  held-out label information and yields optimistic apparent performance.
- **Pipeline B (guarded nested pipeline):** feature selection, scaling, class
  weighting, and hyperparameter tuning happen **only inside training folds**.
  Outer CV gives the honest generalization estimate; inner CV does selection
  and tuning.

This is a **pilot, not the full paper analysis**. It uses GSE25055 only and a
single feature budget (top K = 100). It does **not** use GSE25065, GSE41998,
GSE20194, GSE20271, or any external validation data.

# Cohort and labels

- Accession: **GSE25055** (GPL96 / HG-U133A), discovery cohort.
- Endpoint: **pCR vs RD** from the `pathologic_response_pcr_rd` metadata field.
- Samples coded **NA** are excluded (per the dataset audit, 4 NA-coded samples;
  expected usable counts pCR = 57, RD = 249).

# How to run

```bash
# From the repository root:
Rscript scripts/00_setup.R                       # check packages / session info
Rscript scripts/02_leaky_baseline_gse25055.R     # Pipeline A (leaky)
Rscript scripts/03_nested_smo_svm_gse25055.R     # Pipeline B (guarded nested)
```

Both scripts write per-pipeline metrics and predictions into
`results/pilot_gse25055/`. The chunk below combines them into the pilot
comparison table.

# Pipeline A: naive / leaky baseline

```{r}
#| label: leaky-baseline
source("scripts/02_leaky_baseline_gse25055.R")
leaky <- main()
leaky$metrics
```

# Pipeline B: guarded nested pipeline

```{r}
#| label: nested-pipeline
source("scripts/03_nested_smo_svm_gse25055.R")
nested <- main()
nested$metrics
```

# Side-by-side comparison

We expect the leaky baseline (Pipeline A) to look **more optimistic** than the
guarded nested pipeline (Pipeline B); the gap is the quantity this methodology
paper audits.

```{r}
#| label: comparison
comparison <- rbind(leaky$metrics, nested$metrics)
out_dir <- file.path("tables", "pilot_gse25055")
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)
write.csv(comparison,
          file.path(out_dir, "pilot_performance_comparison.csv"),
          row.names = FALSE)

# Combined predictions for the results directory.
leaky_pred  <- transform(leaky$predictions,  pipeline = "A_leaky_baseline")
nested_pred <- transform(nested$predictions, pipeline = "B_guarded_nested")
common_cols <- intersect(names(leaky_pred), names(nested_pred))
all_pred <- rbind(leaky_pred[, common_cols], nested_pred[, common_cols])
write.csv(all_pred,
          file.path("results", "pilot_gse25055", "pilot_predictions.csv"),
          row.names = FALSE)

comparison
```

# Session info

```{r}
#| label: session-info
writeLines(capture.output(sessionInfo()),
           file.path("results", "pilot_gse25055", "pilot_session_info.txt"))
sessionInfo()
```

# Notes on leakage control

- **Pipeline A is leaky on purpose.** Its t-test feature selection sees every
  sample's label before folds are created. Do not interpret its numbers as a
  generalization estimate.
- **Pipeline B never lets the outer test fold influence fitting.** Feature
  selection and the standardizer are fit on training rows only; SVM cost is
  chosen by inner CV; class imbalance is handled with class weights (no SMOTE
  in this first pilot).
- Random seeds and fold assignments are fixed via `R/00_config.R` (`SEED`).
