---
title: "External Validation: GSE25055 -> GSE25065"
subtitle: "Leakage-aware SMO/SVM omics audit"
format: html
execute:
  echo: true
  warning: false
  message: false
---

## Overview

This notebook reports the first **external-validation** analysis of the study.
The full SMO/SVM pipeline -- t-test feature selection (top K = 100),
standardization, SVM cost, and the decision rule -- was fit and **frozen on
GSE25055 (discovery)** only, then applied **exactly once** to **GSE25065
(external validation)**.

GSE25065 was **not** used for feature selection, hyperparameter tuning,
scaling-parameter estimation, threshold selection, class balancing, or any
model-design decision. GSE25055 and GSE25065 are non-overlapping by design
(the original Hatzis et al. discovery/validation split) and share platform
GPL96.

Pipeline code: `scripts/05_external_validation_gse25065.R`
(run `source(...); main()` to regenerate every table below).

```{r read-outputs}
res_dir <- file.path("results", "external_validation_gse25065")
tab_dir <- file.path("tables",  "external_validation_gse25065")

summary_tab <- read.csv(file.path(tab_dir, "external_validation_summary.csv"))
metrics     <- read.csv(file.path(res_dir, "gse25065_external_metrics.csv"))
preds       <- read.csv(file.path(res_dir, "gse25065_external_predictions.csv"))
feats       <- read.csv(file.path(res_dir, "final_gse25055_selected_features.csv"))
```

## Cohort summary and frozen-model configuration

```{r summary-table}
knitr::kable(summary_tab, col.names = c("metric", "value"))
```

## External-validation metrics (GSE25065)

```{r metrics-table}
knitr::kable(
  data.frame(
    metric = c("AUROC", "PR-AUC", "Balanced accuracy", "MCC",
               "Sensitivity (pCR)", "Specificity (RD)"),
    value  = round(c(metrics$auroc, metrics$pr_auc, metrics$balanced_accuracy,
                     metrics$mcc, metrics$sensitivity, metrics$specificity), 4)
  )
)
```

## Confusion matrix (GSE25065)

```{r confusion}
cm <- table(truth = preds$truth, predicted = preds$predicted)
knitr::kable(as.data.frame.matrix(cm))
```

## Top frozen discovery features

```{r top-features}
knitr::kable(head(feats, 20))
```

## Notes and limitations

The full methodological notes, leakage-control statement, feature-alignment
result, and limitations are in
`results/external_validation_gse25065/external_validation_notes.md`.

Headline: external AUROC (~0.61) is below the discovery guarded nested-CV
estimate (~0.73), the model is biased toward the majority class (high
specificity, low pCR sensitivity), and all 100 frozen probes aligned to GSE25065
(0 missing). This is a leakage-honest transportability estimate from a single
external cohort, not evidence of a deployable biomarker.
