---
title: "Reproducible Omics Evidence Audit (v1)"
subtitle: "GSE25055 discovery / GSE25065 external - leakage-aware SMO/SVM"
format: html
execute:
  echo: true
  warning: false
  message: false
---

## Overview

This notebook presents the first **Reproducible Omics Evidence Audit Table**,
integrating four completed stages -- the dataset audit, the GSE25055
leaky-vs-guarded pilot, the feature-stability analysis, and the GSE25065
external validation -- into one auditable table. **No new modeling is performed
here**; every value is read from previously committed stage outputs and
recombined.

This is a **methodology / audit** result, not a clinical biomarker discovery.

Build script: `scripts/06_build_evidence_audit_table.R`
(run `source(...)` to regenerate the tables below).

```{r read}
audit <- read.csv(file.path("tables", "evidence_audit",
                            "reproducible_omics_evidence_audit.csv"),
                  stringsAsFactors = FALSE, check.names = FALSE)
```

## Evidence audit table

```{r table}
knitr::kable(
  audit,
  col.names = c("Audit domain", "Metric", "Value", "Interpretation", "Manuscript use")
)
```

## Headline numbers

```{r headline}
get_val <- function(m) audit$value[audit$metric == m]
headline <- data.frame(
  Quantity = c("Leakage gap AUROC (leaky - guarded)",
               "Leakage gap PR-AUC (leaky - guarded)",
               "External drop AUROC (nested -> external)",
               "External drop PR-AUC (nested -> external)",
               "External drop sensitivity (nested -> external)",
               "Nogueira feature-stability index"),
  Value = c(get_val("Leakage gap AUROC (leaky - guarded)"),
            get_val("Leakage gap PR-AUC (leaky - guarded)"),
            get_val("Drop AUROC (nested -> external)"),
            get_val("Drop PR-AUC (nested -> external)"),
            get_val("Drop sensitivity (nested -> external)"),
            get_val("Nogueira stability index")),
  stringsAsFactors = FALSE
)
knitr::kable(headline)
```

## Cautious interpretation

- This is a **methodology / audit** result, **not** a clinical biomarker
  discovery; no clinical claims are made.
- The **leaky** pipeline **inflates AUROC/PR-AUC**; the **guarded nested**
  pipeline gives an honest estimate and **improves imbalance-aware metrics**
  (balanced accuracy and MCC).
- **Feature stability is moderate** -- a small reproducible core (28 of 222
  features selected in all folds) coexists with a large unstable tail (102 of
  222 selected only once).
- **External validation shows transportability limits**: discrimination and
  minority-class (pCR) sensitivity fall on the independent GSE25065 cohort,
  while specificity is roughly maintained.

Full narrative, limitations, and unresolved risks:
`results/evidence_audit/evidence_audit_notes.md`.
